%%% Authors are free to create their own special-purpose macros and
%%% environments as required.  Below, a new macro called "\square"
%%% will draw a small box, 5 points high and 5 points wide.  The
%%% environment "finalsquare" is intended for a single paragraph which
%%% ends with a right-justified \square.  If there's no room at the
%%% end of the line, then the square is moved down to the end of the
%%% next line.  Learning to write your own macros can be fun, if
%%% there's a bit of a computer programmer in you. :-)

\newcommand\square{\vbox{\hrule{\hbox{\vrule height5pt\kern5pt\vrule}}\hrule}}
\newenvironment{finalsquare}%
  {\ifvmode\relax\else\par\fi\parfillskip=0pt}%
  {\linebreak[0]\hspace*{\fill}\square\par}

%%% Here begins our chapter.  All heading texts should be capitalized
%%% as if they are titles.  The class file will handle any special
%%% formatting as required.  Note that I use a LaTeX tie between "for"
%%% and "Amortization" here, which prevents LaTeX from inserting a
%%% line break at this location when splitting the chapter title into
%%% multiple lines.  This was a bit of fine-tuning that I added as I
%%% was reviewing the output.  Sometimes even LaTeX needs a little
%%% help. ;-)

\chapter{A Derivation of a Formula for~Amortization}

This chapter contains several examples of the equation environment and
equation references.  There are also examples of every level of
heading, from \verb|\chapter| to \verb|\subparagraph| (though these
headings are somewhat artificial).  The \pkg{amsmath} package is required
to process this chapter, and so the \lit{thesis.tex} file in this
directory contains the package line \verb|\usepackage{amsmath}|.
The text of this file is located in the file \lit{math.tex}.
If you're new to \LaTeX, you may find it instructive to look at the
source text (which contains some extra documentation written as
\LaTeX{} comments) while reviewing this printed output.

The text and mathematics in this document are my own work, written in
the mid-1980s when I was trying to figure out how long it would take
to pay off my credit card debt after I had graduated.  (This work is
the basis of my on-line amortization calculator at the following
web address:
\lit{http://bretwhissel.net/amortization/amortize.html}.)  This
material became a convenient test bed as I developed the FSU thesis
macros.

\section{Definitions}
This is my derivation of the formula for amortization.  The goal is to
find a payment amount, $x$, which pays off the loan principal, $P$,
after a specified number of payments, $N$.  Variable definitions are
listed in table~\ref{tab:varlist}.

%%% The following is an example of a LaTeX table, using the 'tabular'
%%% environment.  It is enclosed within a 'center' environment, which
%%% is also enclosed with the LaTeX 'table' environment.  The 'table'
%%% environment is specified with the placement options '[htpb]'.
%%% Each of these letters indicates a potential placement for this
%%% table.  First, it will be placed [h]ere, if there is room.
%%% Otherwise, the table may float to the [b]ottom of a page, to the
%%% [t]op of a page, or placed on a [p]age by itself.

%%% The 'table' environment can include a '\caption', which should be
%%% placed above the table data.  The \caption is labeled with
%%% '\label{tab:varlist}', which means that I may reference this table
%%% elsewhere in my document.  In fact, the previous paragraph of text
%%% includes a '\ref{tab:varlist}', which will be replaced with the
%%% actual table number when processed by LaTeX. There's a whole
%%% chapter on tables elsewhere in this document

\begin{table}[htbp]
\caption{The List of Variables\label{tab:varlist}}
\begin{center}
\begin{tabular}{cl}
$P$&The principal borrowed\\
$N$&The number of payments\\
$i$&The fractional (periodic) interest rate\\
$P_j$&The principal part of payment $j$\\
$I_j$&The interest part of payment $j$\\
$B$&A final balloon payment\\
$x$&The regular payment
\end{tabular}
\end{center}
\end{table}

%%% Equations can be "inline" or "display" style.  Typically short-ish
%%% equations like $E=mc^2$ may be inlined, or mentions of a single
%%% expression or variable like $x_i$.  More complicated equations are
%%% typically contained in 'equation' or 'align' environments, which
%%% provide a little surrounding vertical space and center equations
%%% in the text column.  Even though these "displayed" equations take
%%% up more space, they should be treated as if they're parts of
%%% sentences, and as such, normal paragraphing rules apply.  I.e., a
%%% blank line before a \begin{equation} will start a new paragraph,
%%% and a blank line following an \end{equation} will end the
%%% paragraph and prepare a new one, which affects white space and
%%% indentation.  Displayed equations usually do not begin
%%% a paragraph, so DON'T leave a blank line before the beginning of
%%% an equation.  Likewise, if you have a line like 'where
%%% x=something...' following a displayed equation, then a blank line
%%% should not precede the 'where...'.

%%% If reference needs to be made to an equation number elsewhere in
%%% the text, then use a \label command in the 'equation' or 'align'
%%% environment so that the label may be referenced later.  There's an
%%% example of a \label command in the equation below.  The required
%%% argument to the \label command is a key, an arbitrary name that
%%% I've assigned to it.

\subsection{Payment Schedule}
Assuming that all payments (excluding an optional final balloon payment)
are the same amount, a payment~$x$ consists of its interest part and
its principal part:
\begin{equation}
\label{amort:components}
x=I_j + P_j
\end{equation}
\begin{align*}
I_1 &= iP               & P_1 &= x - I_1 \\
I_2 &= i(P - P_1)       & P_2 &= x - I_2 \\
I_3 &= i(P - P_1 - P_2) & P_3 &= x - I_3,\mbox{ etc.}
\end{align*}

This schedule states that the payment~$x$ includes interest on all
of the remaining principal, including that which is part of the
current payment.  The first payment, therefore, includes
an interest payment on the total borrowed, which defines the minimum
payment.  (If we are to make any progress toward paying off the loan,
we must pay more than the amount $iP$.)

%%% Learning to type equations in TeX/LaTeX is a bit of an art.  The
%%% TeXbook by Donald Knuth, the creator of TeX, has several examples
%%% and guidelines for typesetting mathematics.  In fact, setting
%%% equations is the reason that TeX was created.  If you have
%%% complicated equations to include in your thesis, you owe it to
%%% yourself to learn more about mathematics typesetting.  Your thesis
%%% is forever, and you want it to look its best. :-) You may find
%%% other references which may be helpful in this regard, such as the
%%% book 'Math into LaTeX' by George Gratzer which provides
%%% comprehensive examples.  You'll probably find many more examples
%%% in on-line resources as well.

The $P_j$'s may be rewritten into a recurrence relation:
\begin{align*}
P_1 &{}= x - iP\\
\noalign{\smallskip}
P_2 &{}= x - i(P - P_1)\\
&{}= x - i\left[P - (x - iP)\right]\\
&{}= x - iP + ix - i^2P\\
&{}= (x - iP)(1 + i)\\
\noalign{\smallskip}
P_3 &{}= x - i(P - P_1 - P_2)\\
&{}= x - i\left[P - (x - iP) - (x - iP + ix - i^2P)\right]\\
&{}= x + 2ix + i^2x - iP - 2i^2P - i^3P\\
&{}= x(1 + i)^2 - iP(1 + i)^2\\
&{}= (x - iP)(1 + i)^2
\end{align*}
In general, we will find that
\begin{equation}
\label{amort:recurrence}
P_j = (x - iP)(1 + i)^{j-1}.
\end{equation}

\section{Balloon Payment}
If there is to be a balloon payment, then the final payment will
consist of the final principal payment $P_f$ and interest on that
principal $i P_f$ so that $B = P_f + iP_f$.  Rewriting $P_f$ in terms
of $B$ gives $P_f = B/(1+i)$.

\section{Finding a Solution}
\subsection{Defining the Equation}
Next, we define an equation which uses these ideas:
\begin{equation}
\label{amort:sums}
B + Nx = P + \sum_{j=1}^N I_j + i\left(\frac{B}{1+i}\right),
\end{equation}
or in English, the sum of all the payments (left side) is equal to the
principal borrowed plus all of the interest paid with regular payments
plus interest paid on the balloon payment (right side).
Note that if there will be no balloon payment ($B=0$), then the $B$
terms drop out.

%%% Note the use of the '\ref' macro below.  There must be
%%% corresponding '\label' commands elsewhere in the document.  The
%%% label/ref keys are arbitrary text.  The keys should probably be
%%% something easy to remember, so that one need not scroll back
%%% through the document to recall what key was used.  All labels
%%% throughout a document must be unique.  A \ref may occur before its
%%% corresponding \label has been defined in the text (i.e., forward
%%% and backward references are allowed).

\subsection{Breaking It Down}
Now we glue some more pieces together: replace $I_j$ of
equation~(\ref{amort:sums}) using the relationship given by
eq.~(\ref{amort:components}) and then substitute the recurrence identity
of eq.~(\ref{amort:recurrence}):
\begin{align}
B - \frac{iB}{1+i} + Nx &{}= 
  P + \sum_{j=1}^N \left[x - (x - iP)(1 + i)^{j-1}\right] \nonumber \\
B - \frac{iB}{1+i} + Nx &{}= 
  P + Nx - (x - iP)\sum_{j=1}^N (1 + i)^{j-1} \nonumber \\
P - B\left(1 - \frac{i}{1+i}\right) &{}= (x - iP)\sum_{j=1}^N (1 + i)^{j-1}
\end{align}

\subsubsection{Initial Solution}
Now we can see our way clear to solve for $x$:
\begin{equation}
\label{amort:series-form}
x = \frac{P - B\left(1-\frac{i}{1+i}\right)}{\displaystyle\sum_{j=1}^{N} (1+i)^{j-1}} 
  + iP.
\end{equation}

\subsubsection{Finding a Closed Form}
While a computer program could be written to solve the problem as
it is, a closed-form solution (i.e., without the iteration) is
preferable.
\paragraph{Isolation} 
The series form of eq.~(\ref{amort:series-form}) can be rewritten
without the series after a little transformation.  First, we separate
the summation and rewrite its limits:
\begin{equation}
x = \left[P - B\left(1 - \frac{i}{1+i}\right)\right]
    \frac{1}{\displaystyle\sum_{j=0}^{N-1}(1+i)^j} + iP.
\end{equation}
\paragraph{Transformation}
To simplify the transformation, we can substitute by letting
$g = 1 + i$ so that the summation looks like $\sum_{j=0}^{N-1} g^j$.  
Next we multiply the series by $(1-g)/(1-g)$, which causes all
but the first and last terms to drop out:
\begin{equation}
\frac{(1-g)\displaystyle\sum_{j=0}^{N-1} g^j}{1-g}=
\frac{\displaystyle\sum_{j=0}^{N-1} g^j - \sum_{j=0}^{N-1} g^{j+1}}{1-g}
  = \frac{1 - g^N}{1-g}.
\end{equation}
\paragraph{Result}
Since the series is originally in the denominator, we invert the transformed
result, and then undo the substitution:
\begin{equation}
\label{amort:closed-form}
x = \left[P-B\left(1-\frac{i}{1+i}\right)\right]\frac{1-(1+i)}{1 - (1+i)^N}
 + iP.
\end{equation}
\paragraph{Rearranged}
Now we can expand and rearrange to taste:
\begin{equation}
\label{amort:rearranged}
x = i \left[\frac{P(1+i)^N}{(1+i)^N - 1} 
  + \frac{B}{(1+i) - (1+i)^{N+1}}\right].
\end{equation}

%%% We created the 'finalsquare' environment at the beginning of this
%%% chapter.  Here's where we use it.

\begin{finalsquare}
\noindent\textit{Quod erat demonstrandum} (``That which was to be
shown''), recognized in most mathematical circles as the initials Q.E.D.
\end{finalsquare}

\section{Application}
\subparagraph{Two forms}
Equations~(\ref{amort:closed-form}) and~(\ref{amort:rearranged})
solve for the payment amount, but either can be re-arranged to solve
for any of the other variables, with the exception of $i$, the
periodic interest rate.  To date I have been unable to find an
analytic solution for this variable, so the program invokes an
iterative method to find successive approximations to the solution.

\subparagraph{Computation}
To reduce the number of computations, $1+i$ can be stored in 
single variable, as well as a single calculation of $(1+i)^N$.
Then the calculation of $(1+i)^{N+1}$ merely requires multiplying the two
previously-calculated values, i.e., $(1+i)^{N+1} = (1+i)^N \cdot (1+i)$.

%%% The chapter ends here.  Actually, it ends wherever the
%%% next \chapter command appears, which is not in this file.  Note
%%% that cross-references using \label, \ref and \pageref aren't
%%% restricted by chapter or file boundaries.  You may make reference
%%% to labels that occur anywhere in your document.
