\chapter{Tabular Data and Tables}

Most graduate students will come to a place in their career where they
must create a table of some kind.  Many simple layouts are a breeze
with \LaTeX.  Here's a brief example:

\begin{quote}
\begin{tabular}{ r r || c | l }
  1  & a      & $3^1$                      &     3 \\
  2  & abb    & $3\times 3^2$              &    27 \\ \hline
  33 & abbccc & $3^3 \times 3^3 \times 3^3$ & 19683 \\
\end{tabular}
\hskip12pt and the source text that created it:
\begin{verbatim}
\begin{tabular}{ r r || c | l }
  1  & a      & $3^1$                       &     3 \\
  2  & abb    & $3\times 3^2$               &    27 \\ \hline
  33 & abbccc & $3^3 \times 3^3 \times 3^3$ & 19683 \\
\end{tabular}
\end{verbatim}
\end{quote}

When the \lit{tabular} environment begins, the next required parameter
specifies the layout of the table.  In this case, the table layout
specifies two right-justified columns (\lit{r r}), a double-vertical
separator (\lit{||}), a centered column (\lit{c}), a single-vertical
separator (\lit{|}), and a left-justified column (\lit{l}).  The next
rows provide the data for the table, with columns separated by the
ampersand (\verb|&|) character.  The end of the row is indicated by
the double backslash (\verb|\\|).  As you can see, the columns may
contain text, numeric data, and even some math.  A horizontal line may
be drawn between rows using the \verb|\hline| command.  As always with
\LaTeX, multiple spaces within columns are ignored.  In addition,
\LaTeX{} also ignores spaces immediately following the \verb|&| character.

For many people, this may be all the information on tables that's
required (for now, anyway).  But at some point, you may need even more
options to create just the right layout.  There are several additional
packages that add functionality to \LaTeX's table-formatting
capability.  A quick web search for \lit{latex table} will turn up a
wealth of usable information, samples, examples, packages, and
tutorials.

The \lit{tabular} environment provides the layout mechanism for
placing text and data into row and column form.  But it's the
\lit{table} environment that allows you to automatically number your
table and to add a heading (caption).  The \lit{table} environment
works just like the \lit{figure} environment as far as floating
placement is concerned.  However, the FSU thesis guidelines state that
table captions should appear \emph{before} the table, while
\lit{figure} captions appear \emph{after} the figure.  The text in
Figure~\ref{sonnet-text} generates Table~\ref{sonnets} as an example.
(And I used \verb|Table~\ref{sonnets}| in the previous sentence to
retrieve the table number.)  This demonstrates how one can create
paragraphs of text as part of a table by using the \verb|p{5cm}|
format specifier.  Additional space was inserted after each row by
adding a dimension to the linebreak specification, i.e., \verb|\\[5pt]|.

However, just because you put some text or data into a multi-column
form, it doesn't necessarily mean that it's a table as far as your
thesis or dissertation is concerned.  If the tabular-form data is part
of your text and flows in the order of your presentation, it may not
be necessary to set it off as a table.  The layout example at the
beginning of this chapter is an example of tabular data which is not
set off as a table.

Other than the unfortunately confusing similarity in their names, the
\lit{tabular} environment and the \lit{table} environment have
independent functionality: while the \lit{tabular} environment is
often used inside the \lit{table} environment, either environment can
be used without the other.  And while we're at it, figures don't
necessarily need to contain graphics.  Figure~\ref{sonnet-text} is an
example of a figure which contains ordinary text, but the text has
been wrapped within a \lit{figure} environment so that it can be
allowed to float outside the main flow of text.

If you have a particularly wide table, you may want to turn the table
sideways on the page.  To do this, add \verb|\usepackage{rotating}| to
the document preamble.  When it is time to insert the rotated table,
type \verb|\begin{sidewaystable}| instead of \verb|\begin{table}|.
This also works for figures, by the way, so instead of
\verb|\begin{figure}|, you may use \verb|\begin{sidewaysfigure}| for
diagrams and images that you want rotated.  Sideways figures and
tables will always be floated to their own page.

\begin{figure}
% Centering doesn't work in a verbatim environment (it takes up the
% entire line-width), so we fake it by prefixing spaces to each line.
\begin{verbatim}
    \begin{table}
    \caption{Shakespeare Sonnets, First Lines, IIX --- XII}
    \label{sonnets}
    \begin{center}
      \begin{tabular}{r p{5cm} }
        8 & Music to hear, why hear'st thou music sadly? \\[5pt]
        9 & Is it for fear to wet a widow's eye \\[5pt]
       10 & For shame deny that thou bear'st love to any \\[5pt]  
       11 & As fast as thou shalt wane, so fast thou grow'st \\[5pt]
       12 & When I do count the clock that tells the time \\
      \end{tabular}
    \end{center}
    \end{table}
\end{verbatim}
\caption{\LaTeX{} source that generates Table~\ref{sonnets} on 
page~\pageref{sonnets}.}
\label{sonnet-text}
\end{figure}

\begin{table}
\caption{Shakespeare Sonnets, First Lines, IIX --- XII}
\label{sonnets}
\begin{center}
\begin{tabular}{r p{5cm} }
  8 & Music to hear, why hear'st thou music sadly? \\[5pt]
  9 & Is it for fear to wet a widow's eye \\[5pt]
 10 & For shame deny that thou bear'st love to any \\[5pt]  
 11 & As fast as thou shalt wane, so fast thou grow'st \\[5pt]
 12 & When I do count the clock that tells the time \\
\end{tabular}
\end{center}
\end{table}
