%%% This chapter is here just to test various features of the FSU
%%% Thesis Class.

\chapter{Testing Appearance}

Today is the first day of the rest of your life.  Birds of a feather
flock together. A narrow mind has a broad tongue.\footnote{It never
  occurred to me to test the footnote function with \texttt{expanded}
  mode line-spacing.  Perhaps it sucks out loud, or perhaps it works
  just fine.  I really need to see it before I can judge. The note
  needs to be long enough that it fills a few lines, just to be sure.}
The reason why worry kills more people than work is that more people
worry than work.  There is more to life than increasing its speed.

\section{No one gets too old\ldots}
\ldots to learn a new way of being stupid.  The journey of a thousand
miles begins with a single step.  Too much of the world runs on the
theory you don't need road manners if you are a five-ton truck.  He
who is known as an early riser can stay in bed till noon.  When you
see a situation you cannot understand, look for the financial
interest.  The following paragraph uses the \lit{singlespaced}
environment.

\begin{singlespaced}
It is better to be a has-been than one who never was.  If Jack's in
love, he's no judge of Jill's beauty.  We didn't all come over on the
same ship, but we're all in the same boat.  After learning the tricks
of the trade, many of us think we know the trade.  Those who cannot
remember the past are condemned to repeat it.
\end{singlespaced}
It is better to be a has-been than one who never was.  If Jack's in
love, he's no judge of Jill's beauty.  We didn't all come over on the
same ship, but we're all in the same boat.  After learning the tricks
of the trade, many of us think we know the trade.  Those who cannot
remember the past are condemned to repeat it.

\section{There is nothing more frightful\ldots}
\ldots than ignorance in action.  He that lives on hope has but a
slender diet.  If God had meant for people to go nude, they would have
been born that way.  Those who deny freedom to others deserve it not
for themselves.  One of these days is none of these days.  There are
two sides to every question and if you want to be popular you take
both.
\begin{quotation}
Logic is the art of going wrong with confidence.  All looks
yellow to a jaundiced eye.  The man who is too old to learn was
probably always too old to learn.  It's what you learn after you know
it all that counts.  The rich need not beg a welcome.  Castles in the
air are all right until we try to move into them.

To most of us it would be very convenient if God were a rascal.  The
only sure thing about luck is that it will change.  It is difficult to
see the picture when you are inside the frame.  We hear and apprehend
only what we already half know.
\end{quotation}
When you probe unseen places use protection.  A real Christian is a
person who can give his pet parrot to the town gossip.  The dog that
trots about finds a bone.  The most welcome person is the one who
knows when to go.  He who loses his head is usually the last one to
miss it.

\subsection{One meets his destiny\ldots}
\ldots often in the road he takes to avoid it.  He that has but four
and spends five has no need for a purse.  Many a live wire would be a
dead one except for his connections.  The vow that binds too strictly
snaps itself.  Life is like playing a violin solo in public and
learning the instrument as one goes on.

\subsubsection{Nothing so needs reforming\ldots}
as other people's habits.  If at first you don't succeed,
you're running about average.  The fellow who is a good sport has to
lose to prove it.  It is easier to fight for one's principles than to
live up to them.  Easy street is a blind alley.

\subsubsection{It would be a swell world\ldots}
if everybody was as pleasant as the fellow who's trying to skin
you.  To say the right thing at the right time, keep still most of the
time.  When it comes to helping you, some people stop at nothing.  The
growth of wisdom can be accurately gauged by the drop in ill temper.

\section{There are visual errors\ldots}
\ldots in time as well as in space.  One thing in favor of a polka-dot
tie is that one more spot doesn't matter.  Occasionally you meet a
person who thinks he is all seven wonders of the world.  You never
know how many friends you have until you rent a cottage at the beach.
Some people think they are generous because they give away free
advice.
\begin{quote}
It is better to risk saving a guilty person than to condemn
an innocent one.

No question is so difficult as that to which the
answer is obvious.

Some people know how to live everyone's life but
their own.  Don't discourage the other man's plans unless you have
better ones to offer.
\end{quote}

The things most people want to know are usually
none of their business.
\begin{verbatim}
10 LOOP I=1 to 10
20 PRINT "Hello, World!"
30 ENDLOOP
40 PRINT "Had enough yet?"
50 GOTO 10
\end{verbatim}
A thing that nobody believes cannot be proved
too often.  Who lives content with little possesses everything.  No
one is unemployed who minds his own business.  Statistics are no
substitute for judgment.

\section{And Again, with Explicit Spaces}
Some people know how to live everyone's life but
their own.  Don't discourage the other man's plans unless you have
better ones to offer.  The things most people want to know are usually
none of their business.
\begin{verbatim*}
  10 LOOP I=1 to 10
  20 PRINT "Hello, World!"
  30 ENDLOOP
  40 PRINT "Had enough yet?"
  50 GOTO 10
\end{verbatim*}
A thing that nobody believes cannot be proved
too often.  Who lives content with little possesses everything.  No
one is unemployed who minds his own business.  Statistics are no
substitute for judgment.

\section{The man who wants\ldots}
\ldots for things to turn up has his eyes fixed on his toes.  The
secret of success in conversation is to be able to disagree without
being disagreeable.  Experience is one thing you can't get for
nothing.  To suspect a friend is worse than to be deceived by him.
\begin{enumerate}
\item The world does not require so much to be informed as to be
  reminded. Knowledge and timber shouldn't be much used till they are
  seasoned.
\item Mistrust first impulses, they are always good.  To approve is
  more difficult than to admire.  The virtue which we appreciate we to
  some extent appropriate.
\begin{itemize}
\item The true worth of a man is to be measured by the objects he pursues.
Be obscure clearly.  Some people pay so much attention to their
reputation that they lose their character.  
\item Anticipation breeds frustration.
\end{itemize}
\item When the cup is full, carry it even.
There is no indigestion
worse than that which comes from having to eat your own words.
\end{enumerate}


Everyone complains of memory, no one of his judgment.  Nothing is
impossible to the man who doesn't have to do it himself.  It may be
those who do most, dream most.  They talk most who have least to say.
A man in passion rides a mad horse.  You cannot be anything if you
want to be everything.  What is said when drunk has been thought out
beforehand.  Have a nice day.
